Binocular combination in the autonomic nervous system

### 2024-07-03 manuscript.tex
Small edits
Add files via upload

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Binocular combination in the autonomic nervous system},
  colorlinks=true,
  linkcolor=black,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption} \usepackage{lineno} \linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Binocular combination in the autonomic nervous system}
\author{Federico G. Segala\(^1\), Aurelio Bruno\(^{1,2}\), Joel T. Martin\(^{1,3}\),\\
Anisa Y. Morsi\(^1\), Alex R. Wade\(^{1,4}\) \& Daniel H. Baker\(^{1,4}\)}
\date{}

@@

Address for correspondence: Federico G. Segala, \url{federico.segala@york.ac.uk}

\hypertarget{abstract-250-words}{%
\section{Abstract (250 words)}\label{abstract-250-words}}

@@

\hypertarget{significance-statement-120-words}{%
\section{Significance statement (120 words)}\label{significance-statement-120-words}}

The diameter of the pupil depends on the level of ambient light and is controlled by distinct retinal photoreceptor classes (cones, rods and intrinsically photosensitive retinal ganglion cells). The computational processes behind binocular integration within the subcortical pathways controlling pupil dimension have been investigated and have been shown to occur independently of visual cortex. We investigated binocular combination within the different photoreceptor pathways using silent substitution. We found different patterns of binocular combination in each pathway and also differences in the phase of the pupil response. The results are consistent with a computational model of binocular signal combination, where the weight of interocular suppression differs across pathways, and demonstrate the algorithm governing binocular combination in the autonomic nervous system.

@@

The autonomic nervous system regulates many involuntary bodily processes, including the constriction and dilation of the pupils in response to light (\protect\hyperlink{ref-McDougal2015}{1}). The anatomical pathway from the retina to the subcortical nuclei controlling the pupillary light response (PLR) is well established: it includes the Pretectal Olivary nucleus (PON), the Superior Cervical ganglion and the Edinger-Westphal nucleus, which project to the iris sphincter muscles that directly control the pupil size (\protect\hyperlink{ref-McDougal2015}{1}--\protect\hyperlink{ref-Wang2015}{3}). Evidence of a binocular component to the PLR is shown by the existence of a consensual response of the pupil (\protect\hyperlink{ref-Wyatt1981}{4}). The anatomical segregation of the subcortical pathway from the rest of the brain means that this binocular combination of signals must occur independently of the cortical processes of binocular integration required for visual perception. Our recent work (\protect\hyperlink{ref-Segala2023}{5}) has shown that the algorithm underlying binocular combination of light in the pupil pathway differs from that in the cortex. Here we extend this paradigm to compare binocular combination in different photoreceptor pathways that feed into the autonomic nervous system.

Different classes of retinal photoreceptors, including cones, rods and melanopsin-containing intrinsically photosensitive retinal ganglion cells (ipRGCs), are directly involved in controlling and maintaining the size of the pupils (\protect\hyperlink{ref-Barrionuevo2014}{6}--\protect\hyperlink{ref-Woelders2018}{12}). The cones have been shown to drive the initial rapid constriction of the pupils (\protect\hyperlink{ref-Mathot2018}{13}), while the slower and longer activation of the ipRGCs shows their role in maintaining the constriction over a prolonged period of time and in regulating the post-illumination pupillary response (\protect\hyperlink{ref-Markwell2010}{14}, \protect\hyperlink{ref-McDougal2010}{15}). The ipRGCs are a unique photoreceptor class that has been discovered recently (\protect\hyperlink{ref-Provencio2000}{16}). They express the photopigment melanopsin, and are involved in the regulation of the circadian rhythm (\protect\hyperlink{ref-Provencio2000}{16}--\protect\hyperlink{ref-Ruby2002}{18}), forming a major input to the PON (\protect\hyperlink{ref-Dacey2003}{19}). The first direct evidence of the involvement of the ipRGCs in the PLR was shown in melanopsin knockout rats by (\protect\hyperlink{ref-Lucas2003}{20}), resulting in the loss of the intrinsic photosensitivity of the cells and a reduced pupil constriction. This same behaviour was later also observed in primates and humans (\protect\hyperlink{ref-Gamlin2007}{21}), where it was demonstrated that the PLR continues during light presentation even when cone and rod signalling is blocked, indicating the primary role of the ipRGCs in maintaining pupil constriction over a prolonged time.

A technique known as silent substitution (\protect\hyperlink{ref-Estevez1982}{22}) can be used to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (\protect\hyperlink{ref-Shapiro1996}{23}, \protect\hyperlink{ref-Spitschan2018}{24}). Studies using silent substitution have shown that the pupils behave differently depending on which photoreceptor or pathway is being stimulated. For example, pupil responses have been shown to be out of phase when the S cones are stimulated, relative to when the ipRGCs or the L+M cones are stimulated (\protect\hyperlink{ref-Spitschan2014}{10}, \protect\hyperlink{ref-Cao2015}{25}). Additionally, when L and M cones are stimulated individually, there appears to be a paradoxical pupil response: increments in L cone stimulation cause a constriction of the pupil, while increments in M cone stimulation lead to a dilation of the pupil (\protect\hyperlink{ref-Murray2018}{9}, \protect\hyperlink{ref-Woelders2018}{12}), perhaps due to activity in the L-M opponent pathway.

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (\protect\hyperlink{ref-Hubel1962}{26}, \protect\hyperlink{ref-Moradi2009}{27}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to only one eye (\protect\hyperlink{ref-Baker2018}{28}--\protect\hyperlink{ref-Legge1984}{30}). At higher contrasts, a process of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (\protect\hyperlink{ref-Moradi2009}{27}, \protect\hyperlink{ref-Baker2007}{31}--\protect\hyperlink{ref-Meese2006}{33}). This is thought to be due to a process of interocular suppression that compensates for the additional excitatory drive caused by stimulating two eyes. Our recent work (\protect\hyperlink{ref-Segala2023}{5}) has shown that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway.

Most studies that use silent substitution to investigate the PLR stimulate only one eye. If the stimulated eye is pharmacologically dilated to control pupil size, the response from the unstimulated eye (known as the consensual response) can be recorded (\protect\hyperlink{ref-Spitschan2014}{10}, \protect\hyperlink{ref-McDougal2010}{15}). While this approach controls the total amount of light incident at the retina, it does not permit investigation of binocular combination for specific photoreceptor pathways. Here we take a different approach, using a custom binocular multiprimary system (see Figure \ref{fig:spectraplots}a,b) that can modulate each (undilated) eye independently, and a binocular eye tracker to measure the diameter of both pupils. We measure the amplitude of pupil modulation (using Fourier analysis) in response to sinusoidal stimulation at 0.5Hz, for stimuli directed at the luminance pathway, L-M cone pathway, S-cone pathway, and melanopsin pathway (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and between monocular stimulation and a dichoptic masking condition in which the two eyes are stimulated at different frequencies (0.5 and 0.4Hz). The results are interpreted using a contemporary model of binocular vision (\protect\hyperlink{ref-Meese2006}{33}) implemented within a hierarchical Bayesian framework.

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation (\protect\hyperlink{ref-Kingdom2015}{34}). We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum (\protect\hyperlink{ref-Busse2009}{35}). The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (where the two eyes received different temporal frequencies; purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

To estimate the extent of interocular suppression for each photoreceptor pathway, we fitted each data set using a Bayesian hierarchical implementation of a simple binocular combination model (\protect\hyperlink{ref-Segala2023}{5}, \protect\hyperlink{ref-Meese2006}{33}). Our primary objective was to compare posterior distributions of the weight of interocular suppression, which are shown in Figure \ref{fig:FirstHarmonicPlots}e. Consistent with our earlier observations, the strongest suppressive weight corresponds to the L-M and Melanopsin conditions, and the weakest suppression corresponds to the luminance and S-cone conditions, with virtually no overlap between the posterior distributions for weak and strong suppression. The model fits were of good quality, as shown by the curves in Figure \ref{fig:FirstHarmonicPlots}a-d,f-i. The fitted model parameters are given in Table \ref{tab:paramtable}.

@@

Finally, we inspected the response phase of our four stimulation conditions, given previous reports that these differ across pathways (\protect\hyperlink{ref-Spitschan2014}{10}), and may be in antiphase for melanopsin and S-cone signals. Figure \ref{fig:phaseplots} shows the phase angles for the first (a) and second (b) harmonic frequencies for binocular stimulation (monocular stimulation produced very similar results). At the first harmonic, melanopsin and S-cone signals differed in phase by more than 90 degrees, though they were not fully in antiphase. The luminance and L-M responses were approximately in antiphase to each other, however this is likely due to our choice to modulate L+M- in the first half-cycle of the sine wave (corresponding to a luminance increase in the luminance condition), and L-M+ in the second half-cycle (corresponding to a luminance decrease in the luminance condition). Had we reversed this phase arrangement, we would likely have seen a close phase correspondence between the L-M and luminance conditions (another way to think about this is that L-cone decreases and M-cone increases are processed like luminance increases). Phase differences between the luminance, S-cone and melanopsin conditions are likely attributable to different lags in phototransduction at the earliest stage (i.e.~in the retina). At the second harmonic frequency (Figure \ref{fig:phaseplots}b), the luminance condition was again out of phase with the other three, and was approximately in quadrature phase with the L-M condition, and in antiphase with both the S-cone and melanopsin conditions (which were in phase with each other).

@@

This is the first study to investigate binocular interactions in the melanopsin pathway directly. A previous meta-analysis (\protect\hyperlink{ref-Spitschan2019}{11}) indicated that there may be a substantial binocular facilitation effect in the circadian pathway, as indexed by melatonin suppression (melatonin is a hormone released by the pineal gland; its production is suppressed by exposure to bright light, particularly when the melanopsin-containing ipRGCs are stimulated). In brief, monocular stimulation of the ipRGCs requires up to ten times the signal strength to produce an equivalent effect to binocular stimulation. This superadditive effect implies an absence of interocular suppression, and perhaps the presence of either a highly compressive nonlinearity, or an AND-style neural operation. Our findings here are very different - we do not see substantial binocular facilitation effects in response to melanopsin modulation, and our data indicate that interocular suppression is very strong. Moreover, by measuring the full contrast response function, we can rule out compressive nonlinearities, because the functions accelerate at both the first and second harmonic frequencies. The explanation for this difference could be due to different anatomical pathways. Binocular combination in the circadian system likely takes place in the suprachiasmatic nucleus, whereas in the pupil constriction circuit the Edinger-Westphal nucleus is the most likely site of binocular integration (\protect\hyperlink{ref-Mathot2018}{13}). Presumably these anatomical differences, and the practical constraints of the two systems, lead to differences in response.

Our recent psychophysical work (\protect\hyperlink{ref-Baker2024}{36}) has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the luminance pathway. Using a contrast discrimination paradigm with spatial modulations of luminance and colour, (\protect\hyperlink{ref-Baker2024}{36}) found equally strong interocular suppression in all three pathways. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the luminance and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in (\protect\hyperlink{ref-Baker2024}{36}). Our other recent work (\protect\hyperlink{ref-Segala2023}{5}) has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations (\protect\hyperlink{ref-Baker2017}{37}). These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' (\protect\hyperlink{ref-Baker2007}{31}), in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes (\protect\hyperlink{ref-Kaestner2019}{38}). Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

@@

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

@@

Before the start of the L-M experiment, participants completed a luminance nulling perceptual calibration procedure in L-M cone space on an Iiyama \(\textrm{VisionMaster}^{\textrm{TM}}\) Pro 510 display (800 x 600 pixels, 60 Hz refresh rate). During the task, participants were presented with a disc flickering within the L-M cone space (between magenta and cyan). Using a trackball, participants adjusted the angle in cone space to find their subjective isoluminant point, which resulted in changing the flickering intensity of the stimulus until the amplitude of the flicker appeared to be minimised. The result was used to modify the requested contrasts during stimulus preparation so as to account for individual differences affecting perceived illuminance, principally the L:M cone ratio (\protect\hyperlink{ref-Carroll2002}{49}, \protect\hyperlink{ref-Hofer2005}{50}).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Binocular combination in the autonomic nervous system},
  colorlinks=true,
  linkcolor=black,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Binocular combination in the autonomic nervous system}
\author{Federico G. Segala\(^1\)*, Aurelio Bruno\(^{1,2}\), Joel T. Martin\(^{1,3}\),\\
Anisa Y. Morsi\(^1\), Alex R. Wade\(^{1,4}\) \& Daniel H. Baker\(^{1,4}\)}
\date{}

@@

*Corresponding author

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

@@

@@

The autonomic nervous system regulates many involuntary bodily processes, including the constriction and dilation of the pupils in response to light.\textsuperscript{\protect\hyperlink{ref-McDougal2015}{1}} The anatomical pathway from the retina to the subcortical nuclei controlling the pupillary light response (PLR) is well established: it includes the Pretectal Olivary nucleus (PON), the Superior Cervical ganglion and the Edinger-Westphal nucleus, which project to the iris sphincter muscles that directly control the pupil size.\textsuperscript{\protect\hyperlink{ref-McDougal2015}{1}--\protect\hyperlink{ref-Wang2015}{3}} Evidence of a binocular component to the PLR is shown by the existence of a consensual response of the pupil (the stimulation of one eye will cause constriction of the other eye)\textsuperscript{\protect\hyperlink{ref-Wyatt1981}{4}}. The anatomical segregation of the subcortical pathway from the rest of the brain means that this binocular combination of signals must occur independently of the cortical processes of binocular integration required for visual perception. Our recent work\textsuperscript{\protect\hyperlink{ref-Segala2023}{5}} has shown that the algorithm underlying binocular combination of light in the pupil pathway differs from that in the cortex. Here we extend this paradigm to compare binocular combination in different photoreceptor pathways that feed into the autonomic nervous system.

Different classes of retinal photoreceptors, including cones, rods and melanopsin-containing intrinsically photosensitive retinal ganglion cells (ipRGCs), are directly involved in controlling and maintaining the size of the pupils.\textsuperscript{\protect\hyperlink{ref-Barrionuevo2014}{6}--\protect\hyperlink{ref-Woelders2018}{12}} The cones have been shown to drive the initial rapid constriction of the pupils,\textsuperscript{\protect\hyperlink{ref-Mathot2018}{13}} while the slower and longer activation of the ipRGCs shows their role in maintaining the constriction over a prolonged period of time and in regulating the post-illumination pupillary response.\textsuperscript{\protect\hyperlink{ref-Markwell2010}{14},\protect\hyperlink{ref-McDougal2010}{15}} The ipRGCs are a unique photoreceptor class that has been discovered recently.\textsuperscript{\protect\hyperlink{ref-Provencio2000}{16}} They express the photopigment melanopsin, and are involved in the regulation of the circadian rhythm,\textsuperscript{\protect\hyperlink{ref-Provencio2000}{16}--\protect\hyperlink{ref-Ruby2002}{18}} forming a major input to the PON.\textsuperscript{\protect\hyperlink{ref-Dacey2003}{19}} The first direct evidence of the involvement of the ipRGCs in the PLR was shown in melanopsin knockout rats by Lucas et al.~\textsuperscript{\protect\hyperlink{ref-Lucas2003}{20}}, resulting in the loss of the intrinsic photosensitivity of the cells and a reduced pupil constriction. This same behaviour was later also observed in primates and humans,\textsuperscript{\protect\hyperlink{ref-Gamlin2007}{21}} where it was demonstrated that the PLR continues during light presentation even when cone and rod signalling is blocked, indicating the primary role of the ipRGCs in maintaining pupil constriction over a prolonged time.

A technique known as silent substitution\textsuperscript{\protect\hyperlink{ref-Estevez1982}{22}} can be used to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them.\textsuperscript{\protect\hyperlink{ref-Shapiro1996}{23},\protect\hyperlink{ref-Spitschan2018}{24}} Studies using silent substitution have shown that the pupils behave differently depending on which photoreceptor or pathway is being stimulated. For example, pupil responses have been shown to be out of phase when the S cones are stimulated, relative to when the ipRGCs or the L+M cones are stimulated.\textsuperscript{\protect\hyperlink{ref-Spitschan2014}{10},\protect\hyperlink{ref-Cao2015}{25}} Additionally, when L and M cones are stimulated individually, there appears to be a paradoxical pupil response: increments in L cone stimulation cause a constriction of the pupil, while increments in M cone stimulation lead to a dilation of the pupil,\textsuperscript{\protect\hyperlink{ref-Murray2018}{9},\protect\hyperlink{ref-Woelders2018}{12}} perhaps due to activity in the L-M opponent pathway.

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex.\textsuperscript{\protect\hyperlink{ref-Hubel1962}{26},\protect\hyperlink{ref-Moradi2009}{27}} For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to only one eye.\textsuperscript{\protect\hyperlink{ref-Baker2018}{28}--\protect\hyperlink{ref-Legge1984}{30}} At higher contrasts, a process of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal.\textsuperscript{\protect\hyperlink{ref-Moradi2009}{27},\protect\hyperlink{ref-Baker2007}{31}--\protect\hyperlink{ref-Meese2006}{33}} This is thought to be due to a process of interocular suppression that compensates for the additional excitatory drive caused by stimulating two eyes. Our recent work\textsuperscript{\protect\hyperlink{ref-Segala2023}{5}} has shown that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway.

Most studies that use silent substitution to investigate the PLR stimulate only one eye. If the stimulated eye is pharmacologically dilated to control pupil size, the response from the unstimulated eye (known as the consensual response) can be recorded.\textsuperscript{\protect\hyperlink{ref-Spitschan2014}{10},\protect\hyperlink{ref-McDougal2010}{15}} While this approach controls the total amount of light incident at the retina, it does not permit investigation of binocular combination for specific photoreceptor pathways. Here we take a different approach, using a custom binocular multiprimary system (see Figure \ref{fig:spectraplots}a,b) that can modulate each (undilated) eye independently, and a binocular eye tracker to measure the diameter of both pupils. We measure the amplitude of pupil modulation (using Fourier analysis) in response to sinusoidal stimulation at 0.5Hz, for stimuli directed at the luminance pathway, L-M cone pathway, S-cone pathway, and melanopsin pathway (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and between monocular stimulation and a dichoptic masking condition in which the two eyes are stimulated at different frequencies (0.5 and 0.4Hz). The results are interpreted using a contemporary model of binocular vision\textsuperscript{\protect\hyperlink{ref-Meese2006}{33}} implemented within a hierarchical Bayesian framework.

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation.\textsuperscript{\protect\hyperlink{ref-Kingdom2015}{34}} We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum.\textsuperscript{\protect\hyperlink{ref-Busse2009}{35}} The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (where the two eyes received different temporal frequencies; purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

To estimate the extent of interocular suppression for each photoreceptor pathway, we fitted each data set using a Bayesian hierarchical implementation of a simple binocular combination model.\textsuperscript{\protect\hyperlink{ref-Segala2023}{5},\protect\hyperlink{ref-Meese2006}{33}} Our primary objective was to compare posterior distributions of the weight of interocular suppression, which are shown in Figure \ref{fig:FirstHarmonicPlots}e. Consistent with our earlier observations, the strongest suppressive weight corresponds to the L-M and Melanopsin conditions, and the weakest suppression corresponds to the luminance and S-cone conditions, with virtually no overlap between the posterior distributions for weak and strong suppression. The model fits were of good quality, as shown by the curves in Figure \ref{fig:FirstHarmonicPlots}a-d,f-i. The fitted model parameters are given in Table \ref{tab:paramtable}.

@@

Finally, we inspected the response phase of our four stimulation conditions, given previous reports that these differ across pathways,\textsuperscript{\protect\hyperlink{ref-Spitschan2014}{10}} and may be in antiphase for melanopsin and S-cone signals. Figure \ref{fig:phaseplots} shows the phase angles for the first (a) and second (b) harmonic frequencies for binocular stimulation (monocular stimulation produced very similar results). At the first harmonic, melanopsin and S-cone signals differed in phase by more than 90 degrees, though they were not fully in antiphase. The luminance and L-M responses were approximately in antiphase to each other, however this is likely due to our choice to modulate L+M- in the first half-cycle of the sine wave (corresponding to a luminance increase in the luminance condition), and L-M+ in the second half-cycle (corresponding to a luminance decrease in the luminance condition). Had we reversed this phase arrangement, we would likely have seen a close phase correspondence between the L-M and luminance conditions (another way to think about this is that L-cone decreases and M-cone increases are processed like luminance increases). Phase differences between the luminance, S-cone and melanopsin conditions are likely attributable to different lags in phototransduction at the earliest stage (i.e.~in the retina). At the second harmonic frequency (Figure \ref{fig:phaseplots}b), the luminance condition was again out of phase with the other three, and was approximately in quadrature phase with the L-M condition, and in antiphase with both the S-cone and melanopsin conditions (which were in phase with each other).

@@

This is the first study to investigate binocular interactions in the melanopsin pathway directly. A previous meta-analysis\textsuperscript{\protect\hyperlink{ref-Spitschan2019}{11}} indicated that there may be a substantial binocular facilitation effect in the circadian pathway, as indexed by melatonin suppression (melatonin is a hormone released by the pineal gland; its production is suppressed by exposure to bright light, particularly when the melanopsin-containing ipRGCs are stimulated). In brief, monocular stimulation of the ipRGCs requires up to ten times the signal strength to produce an equivalent effect to binocular stimulation. This superadditive effect implies an absence of interocular suppression, and perhaps the presence of either a highly compressive nonlinearity, or an AND-style neural operation. Our findings here are very different - we do not see substantial binocular facilitation effects in response to melanopsin modulation, and our data indicate that interocular suppression is very strong. Moreover, by measuring the full contrast response function, we can rule out compressive nonlinearities, because the functions accelerate at both the first and second harmonic frequencies. The explanation for this difference could be due to different anatomical pathways. Binocular combination in the circadian system likely takes place in the suprachiasmatic nucleus, whereas in the pupil constriction circuit the Edinger-Westphal nucleus is the most likely site of binocular integration.\textsuperscript{\protect\hyperlink{ref-Mathot2018}{13}} Presumably these anatomical differences, and the practical constraints of the two systems, lead to differences in response.

Our recent psychophysical work\textsuperscript{\protect\hyperlink{ref-Baker2024}{36}} has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the luminance pathway. Using a contrast discrimination paradigm with spatial modulations of luminance and colour, we found equally strong interocular suppression in all three pathways. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the luminance and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in our previous work.\textsuperscript{\protect\hyperlink{ref-Baker2024}{36}} Our other recent work\textsuperscript{\protect\hyperlink{ref-Segala2023}{5}} has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations.\textsuperscript{\protect\hyperlink{ref-Baker2017}{37}} These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance',\textsuperscript{\protect\hyperlink{ref-Baker2007}{31}} in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes.\textsuperscript{\protect\hyperlink{ref-Kaestner2019}{38}} Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

Before the start of the L-M experiment, participants completed a luminance nulling perceptual calibration procedure in L-M cone space on an Iiyama \(\textrm{VisionMaster}^{\textrm{TM}}\) Pro 510 display (800 x 600 pixels, 60 Hz refresh rate). During the task, participants were presented with a disc flickering within the L-M cone space (between magenta and cyan). Using a trackball, participants adjusted the angle in cone space to find their subjective isoluminant point, which resulted in changing the flickering intensity of the stimulus until the amplitude of the flicker appeared to be minimised. The result was used to modify the requested contrasts during stimulus preparation so as to account for individual differences affecting perceived illuminance, principally the L:M cone ratio.\textsuperscript{\protect\hyperlink{ref-Carroll2002}{49},\protect\hyperlink{ref-Hofer2005}{50}}
